QPLIB: A Library of Quadratic Programming Instances

### 2016-12-27 format.tex
Update format.tex

The QPLIB format is designed to cope with a variety of
linear and quadratic programs. We distinguish, in particular,
linear programs (LP)
\begin{align*}
 \mininx{} & l^T x + f \\
 \st & c\s{L} \leq A x \leq c\s{U} \tim{and} x\s{L} \leq x \leq x\s{U}, \\
\intertext{linear programs with quadratic constraints (LPQC)}
 \mininx{} & l^T x + f \\
 \st & c\s{L} \leq A x + \half  [x,Q\s{C},x] \leq c\s{U}
     \tim{and} x\s{L} \leq x \leq x\s{U}, \\
\intertext{bound-constrained quadratic programs (BQP)}
 \mininx{} & \half x^T Q x + l^T x + f \\
 \st & x\s{L} \leq x \leq x\s{U}, \\
\intertext{quadratic programs (QP)}
 \mininx{} & \half x^T Q x + l^T x + f \\
 \st & c\s{L} \leq A x \leq c\s{U} \tim{and} x\s{L} \leq x \leq x\s{U}, \\
\intertext{and quadratic program with quadratic constraints (QPQC)}
 \mininx{} & \half x^T Q x + l^T x + f \\
 \st & c\s{L} \leq A x + \half [x,Q\s{C},x] \leq c\s{U}
     \tim{and} x\s{L} \leq x \leq x\s{U},
\end{align*}
\todo[inline]{\small Stefan: Are LPQC and QPQC frequently used abbreviations? If a linear program has quadratic constraints, how can it then still be a linear program? I am more familiar with the abbreviations QCP instead of LPQC and QCQP instead of QPQC. I further think that one could skip QCP and consider them as part of QCQP.}
where $Q\s{C}$ is the diagonal matrix with blocks $Q\s{C}_k$, $k=1,\ldots,m$,
$Q$ and the $Q\s{C}_k$ are symmetric $n$ by $n$ Hessian matrices,
$A$ is a rectangular $m$ by $n$ constraint matrix,
$l$, $x\s{L}$, $x\s{U}$, $c\s{L}$, $c\s{U}$
are vectors of appropriate sizes and $f$ is a scalar; the notation
$[ x, Q\s{C},x]$ denotes the vector whose $k$-th component is
$x^T Q\s{C}_k x$.
In addition, all or a subset of the variables $x$ may be required to be
integer, while further restrictions (such as a variable being binary)
may be imposed via explicit constraints.

@@

$m$  & number of general linear constraints (integer) & [2] \\

@@

$n\s{Q}$ & number of nonzeros  (integer) in lower triangle of $Q$  & [3] \\
$i$\; $j$\; $Q_{i,j}$ & \lbox{row and column indices (integers) and value (real)
for each nonzero entry of $Q$, if $n\s{Q} > 0$, one triple on each line} & \\

@@

$l_d$ & default value (real) for entries in $l$ & \\
$n\s{l}$ & number of non-default entries (integer) in $l$ & \\
$i$\; $l_i$ & \lbox{index (integer) and value (real) for each non-default
term in $l$, if $n\s{l} > 0$, one pair per line} & \\

@@

$f$ & value (real) of f & \\ \hline
$n\s{Q${}_c$}$ & number of nonzeros (integer) in lower triangle of
$Q\s{C}$  & [2,4] \\
$k$\; $i$\; $j$\; $Q\s{C}_{k,i,j}$

@@

 each entry of $Q\s{C}$, if $n\s{Q${}_c$} > 0$, one quadruple on each line} & \\

@@

$n\s{A}$ & number of nonzeros  (integer) in $A$                     & [2] \\
$i$\; $j$\; $A_{i,j}$ & \lbox{row and column indices (integers) and value (real)

@@

entry of $A$, if $n\s{A} > 0$, one triple on each line} & \\

@@

$c\s{L}_d$ & default value (real) for entries in $c\s{L}$            & [2] \\
$n^{c\s{L}}$ & number of non-default entries (integer) in $c\s{L}$    & [2] \\
$i$\;\ $c\s{L}_i$ & \lbox{index (integer) and value (real) for each non-default

@@

in $c\s{L}$, if $n^{c\s{L}} > 0$, one pair per line}        & [2] \\

@@

$c\s{U}_d$ & default value (real) for entries in $c\s{U}$            & [2] \\
$n^{c\s{U}}$ & number of non-default entries (integer) in $c\s{U}$    & [2] \\
$i$\;\ $c\s{U}_i$ & \lbox{index (integer) and value (real) for each non-default

@@

in $c\s{U}$, if $n^{c\s{U}} > 0$, one pair per line}        & [2] \\

@@

$x\s{L}_d$ & default value (real) for entries in $x\s{L}$            &     \\
$n^{x\s{L}}$ & number of non-default entries (integer) in $x\s{L}$    &     \\
$i$\;\ $x\s{L}_i$ & \lbox{index (integer) and value (real) for each non-default
term in $x\s{L}$, if $n^{x\s{L}} > 0$, one pair per line} & \\

@@

$x\s{U}_d$ & default value (real) for entries in $x\s{U}$            &     \\
$n^{x\s{U}}$ & number of non-default entries (integer) in $x\s{U}$    &     \\
$i$\;\ $x\s{U}_i$ & \lbox{index (integer) and value (real) for each non-default
term in $x\s{U}$, if $n^{x\s{U}} > 0$, one pair per line} & \\

@@

$i$\; $x_i$ & \lbox{index (integer) and name (character string)

@@

$i$\; $c_i$ & \lbox{index  (integer) and name (character string) for each

@@

\begin{description}
\item [{[1]}]
The problem type is represented by a character string as one of the following:
\begin{description}
\item For continuous problems (i.e., all variables are continuous):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [LP]      a linear program,
\item [LPQC]    a linear program with quadratic constraints,
\item [BQP]     a bound-constrained quadratic program,
\item [QP]      a quadratic program, and
\item [QPQC]    a quadratic program with quadratic constraints.
\end{description}
\item For integer problems (i.e., all variables are integer valued):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [ILP]     an integer linear program,
\item [ILPQC]   an integer linear program with quadratic constraints,
\item [IBQP]    an integer bound-constrained quadratic program,
\item [IQP]     an integer quadratic program, and
\item [IQPQC]   an integer quadratic program with quadratic constraints.
\end{description}
\item For mixed-integer problems (i.e., there is a mix of continuous and
integer variables):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [MILP]    a mixed-integer linear program,
\item [MILPQC]  a mixed-integer linear program with quadratic constraints,
\item [MIBQP]   a mixed-integer bound-constrained quadratic program,
\item [MIQP]    a mixed-integer quadratic program, and
\item [MIQPQC]  a mixed-integer quadratic program with quadratic constraints.
\end{description}
\todo[inline]{\small Stefan: I agree that leaving out information on
numerical properties of the matrices and vectors (convexity, etc.) is good here.
Such properties may also depend on tolerances, etc.
As above, I would suggest to eliminate LPQC and rename QPQC into QCP.
Having a string for ``unspecified'' may be nice, too, or one just uses MIQCQP in this case.}
\end{description}

@@

As a simple example, consider the mixed-integer QP
\[\arr{l}{\mininx{} x_1^2 + x_2^2 + x_3^2 - x_1 x_2 - x_2 x_3
  - 0.2 x_1  -0.4 x_2 - 0.2 x_3 \\
\st 1 \leq x_1 + x_2, \; 1 \leq x_1 + x_3, \; 0 \leq x_1 \leq 1,
\; 0 \leq x_2 \leq 2, \; \mbox{and binary} \; x_3.}
\]
This may then be represent in QPLIB format as follows:

{\small
\begin{verbatim}
! ---------------
! example problem
! ---------------
MIPBAND   # problem name
MIQP      # problem is a mixed-integer quadratic program
3         # variables
2         # general linear constraints
5         # nonzeros in lower triangle of Q
1 1 2.0   5 lines of row & column index & value of nonzeros in lower triangle Q
2 1 -1.0  |
2 2 2.0   |
3 2 -1.0  |
3 3 2.0   |
-0.2      default value for entries in l
1         # non default entries in l
2 -0.4    1 lines of index & value of non-default values in l
0.0       value of f
4         # nonzeros in A
1 1 1.0   4 lines of row & column index & value of nonzeros in A
1 2 1.0   |
2 1 1.0   |
2 3 1.0   |
1.0E+20   infinity
1.0       default value for entries in c^L
0         # non default entries in c^L
1.0E+20   default value for entries in c^U
0         # non default entries in c^U
0.0       default value for entries in x^L
0         # non default entries in x^L
1.0       default value for entries in x^U
1         # non default entries in x^U
2 2.0     1 line of non-default indices and values in x^U
0         default variable type is continuous
1         # non default variable types
3 1       variable 3 is integer
1.0       default value for initial values for x
0         # non default entries in x
0.0       default value for initial values for y
0         # non default entries in y
0.0       default value for initial values for z
0         # non default entries in z
0         # non default names for variables
0         # non default names for constraints
\end{verbatim}
}

>>>>

Following the notation of Section~\ref{sec:QPbasic}, a generic Quadratic Program (QP) can be written as follows:
%
\begin{align*}	
 \min \;\;
 & x^\top Q^0 x + b^0 x + f \\
 %
 & c^i_l \leq x^\top Q^i x + b^i x \leq c^i_u & i \in \Mcm \\ 
 %
 & l_j \leq x_j \leq u_j & j \in \Mcn  \\
 %
 & x_j \in \mathbb{Z} & j \in \Mcz
\end{align*}

where:

\begin{itemize}
 \item $\Mcn = \{ 1, \ldots, n \}$ is the set of (indices) of variables, and $\Mcm = \{ 1, \ldots, m \}$ is the set of (indices) of constraints;
 %
 \item $x = [x_j]_{j = 1}^n \in \RR^n$ is a finite vector of real variables;
 %
 \item $Q^i$ for $i \in \{ 0 \} \cup \Mcm$ are symmetric $n \times n$ real matrices;
 %
 \item $b^i$, $c^i_l$, $c^i_u$ for $i \in \{ 0 \} \cup \Mcm$ and $f$ are, respectively, real $n$-vectors, real constants, real constants and real constant;
 %
 \item $-\infty \leq l_j \leq u_j \leq \infty$ are the (extended) real upper and lower bounds on each variable $x_j$ for $j \in \Mcn$;
 %
 \item $\Mcm = \Mcq \cup \Mcl$ where $Q^i = 0$ for all $i \in \Mcl$;
 %
 \item the variables in $\Mcz \subseteq \Mcm$ are restricted to only attain integer values.
\end{itemize}

%Accordingly, the QPLIB format is designed to cope with a variety of linear and quadratic programs. 
%We distinguish, in particular,
%linear programs (LP)
%\begin{align*}
% \mininx{} & l^T x + f \\
% \st & c\s{L} \leq A x \leq c\s{U} \tim{and} x\s{L} \leq x \leq x\s{U}, \\
%\intertext{linear programs with quadratic constraints (LPQC)}
% \mininx{} & l^T x + f \\
% \st & c\s{L} \leq A x + \half  [x,Q\s{C},x] \leq c\s{U}
%     \tim{and} x\s{L} \leq x \leq x\s{U}, \\
%\intertext{bound-constrained quadratic programs (BQP)}
% \mininx{} & \half x^T Q x + l^T x + f \\
% \st & x\s{L} \leq x \leq x\s{U}, \\
%\intertext{quadratic programs (QP)}
% \mininx{} & \half x^T Q x + l^T x + f \\
% \st & c\s{L} \leq A x \leq c\s{U} \tim{and} x\s{L} \leq x \leq x\s{U}, \\
%\intertext{and quadratic program with quadratic constraints (QPQC)}
% \mininx{} & \half x^T Q x + l^T x + f \\
% \st & c\s{L} \leq A x + \half [x,Q\s{C},x] \leq c\s{U}
%     \tim{and} x\s{L} \leq x \leq x\s{U},
%\end{align*}
%\todo[inline]{\small Stefan: Are LPQC and QPQC frequently used abbreviations? If a linear program has quadratic constraints, how can it then still be a linear program? I am more familiar with the abbreviations QCP instead of LPQC and QCQP instead of QPQC. I further think that one could skip QCP and consider them as part of QCQP.}
%where $Q\s{C}$ is the diagonal matrix with blocks $Q\s{C}_k$, $k=1,\ldots,m$,
%$Q$ and the $Q\s{C}_k$ are symmetric $n$ by $n$ Hessian matrices,
%$A$ is a rectangular $m$ by $n$ constraint matrix,
%$l$, $x\s{L}$, $x\s{U}$, $c\s{L}$, $c\s{U}$
%are vectors of appropriate sizes and $f$ is a scalar; the notation
%$[ x, Q\s{C},x]$ denotes the vector whose $k$-th component is
%$x^T Q\s{C}_k x$.
%In addition, all or a subset of the variables $x$ may be required to be
%integer, while further restrictions (such as a variable being binary)
%may be imposed via explicit constraints.

@@

$m$  & number of constraints (integer) & [2] \\

@@

$n^{Q^0}$ & number of nonzeros  (integer) in lower triangle of $Q^0$  & [3] \\
$h$\; $k$\; $Q_{hk}^0$ & \lbox{row and column indices (integers) and value (real)
for each nonzero entry of $Q^0$, if $n^{Q^0} > 0$, one triple on each line} & \\

@@

$b^0_d$ & default value (real) for entries in $b^0$ & \\
$n^{b^0}$ & number of non-default entries (integer) in $b^0$ & \\
$j$\; $b^0_j$ & \lbox{index (integer) and value (real) for each non-default
term in $b^0$, if $n^{b^0} > 0$, one pair per line} & \\

@@

$f$ & constant part of the objective function & \\ \hline
$n^{Q^i}$ & number of nonzeros (integer) in lower triangle of
$Q^i$  & [2,4] \\
$i$\; $h$\; $k$\; $Q^{i}_{hk}$

@@

 each entry of $Q^{i}$, if $n^{Q^i}$, one quadruple on each line} & \\

@@

$n^{b}$ & number of nonzeros  (integer) in $b^i$, $i \in \Mcm$                     & [2] \\
$i$\; $j$\; $b^i_{j}$ & \lbox{row and column indices (integers) and value (real)

@@

entry of $b^i$, if $n^{b} > 0$, one triple on each line} & \\

@@

$c_{l,d}$ & default value (real) for entries in $c_{l}$            & [2] \\
$n^{c_{l,d}}$ & number of non-default entries (integer) in $c_{l}$    & [2] \\
$i$\;\ $c^i_{l}$ & \lbox{index (integer) and value (real) for each non-default

@@

in $c_{l,d}$, if $n^{c_{l,d}} > 0$, one pair per line}        & [2] \\

@@

$c_{u,d}$ & default value (real) for entries in $c_{u}$            & [2] \\
$n^{c_{u,d}}$ & number of non-default entries (integer) in $c_{u}$    & [2] \\
$i$\;\ $c^i_{u}$ & \lbox{index (integer) and value (real) for each non-default

@@

in $c_{u,d}$, if $n^{c_{u,d}} > 0$, one pair per line}        & \\

@@

$l_{d}$ & default value (real) for entries in $l$            & \\
$n^{l_{d}}$ & number of non-default entries (integer) in $l$    &  \\
$i$\;\ $l_{i}$ & \lbox{index (integer) and value (real) for each non-default
in $l$, if $n^{l_{d}}> 0$, one pair per line}        &  \\

@@

$u_{d}$ & default value (real) for entries in $u$            & \\
$n^{u_{d}}$ & number of non-default entries (integer) in $u$    &  \\
$i$\;\ $u_{i}$ & \lbox{index (integer) and value (real) for each non-default
in $u$, if $n^{u_{d}}> 0$, one pair per line}        &  \\

@@

$j$\; ${\tt var\_name}_j$ & \lbox{index (integer) and name (character string)

@@

$i$\; ${\tt cons\_name}_i$ & \lbox{index  (integer) and name (character string) for each

@@

\begin{description}
\item [{[1]}]
\todo[inline]{\small Fabio and Emiliano: we think that, if possible, we should follow the classification proposed in Section 2.}

The problem type is represented by a character string as one of the following:
\begin{description}
\item For continuous problems (i.e., all variables are continuous):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [LP]      a linear program,
\item [LPQC]    a linear program with quadratic constraints,
\item [BQP]     a bound-constrained quadratic program,
\item [QP]      a quadratic program, and
\item [QPQC]    a quadratic program with quadratic constraints.
\end{description}
\item For integer problems (i.e., all variables are integer valued):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [ILP]     an integer linear program,
\item [ILPQC]   an integer linear program with quadratic constraints,
\item [IBQP]    an integer bound-constrained quadratic program,
\item [IQP]     an integer quadratic program, and
\item [IQPQC]   an integer quadratic program with quadratic constraints.
\end{description}
\item For mixed-integer problems (i.e., there is a mix of continuous and
integer variables):
\begin{description}[leftmargin=!,labelwidth=\widthof{\ttfamily SMILQPC}]
\item [MILP]    a mixed-integer linear program,
\item [MILPQC]  a mixed-integer linear program with quadratic constraints,
\item [MIBQP]   a mixed-integer bound-constrained quadratic program,
\item [MIQP]    a mixed-integer quadratic program, and
\item [MIQPQC]  a mixed-integer quadratic program with quadratic constraints.
\end{description}
\todo[inline]{\small Stefan: I agree that leaving out information on
numerical properties of the matrices and vectors (convexity, etc.) is good here.
Such properties may also depend on tolerances, etc.
As above, I would suggest to eliminate LPQC and rename QPQC into QCP.
Having a string for ``unspecified'' may be nice, too, or one just uses MIQCQP in this case.}
\end{description}

@@

As a simple example, consider the mixed-integer QP
\[\arr{l}{\min_{x \in \smallRe^3} x_1^2 + x_2^2 + x_3^2 - x_1 x_2 - x_2 x_3
  - 0.2 x_1  -0.4 x_2 - 0.2 x_3 \\
\st 1 \leq x_1 + x_2, \; 1 \leq x_1 + x_3, \; 0 \leq x_1 \leq 1,
\; 0 \leq x_2 \leq 2, \; \mbox{and binary} \; x_3.}
\]
This may then be represent in QPLIB format as follows:

{\small
\begin{verbatim}
! ---------------
! example problem
! ---------------
MIPBAND   # problem name
MIQP      # problem is a mixed-integer quadratic program
3         # variables
2         # general linear constraints
5         # nonzeros in lower triangle of Q^0
1 1 2.0   5 lines of row & column index & value of nonzeros in lower triangle Q^0
2 1 -1.0  |
2 2 2.0   |
3 2 -1.0  |
3 3 2.0   |
-0.2      default value for entries in b_0
1         # non default entries in b_0
2 -0.4    1 lines of index & value of non-default values in b_0
0.0       value of f
4         # nonzeros in vectors b^i (i =0,...,m)
1 1 1.0   4 lines of row & column index & value of nonzeros in b^i (i =0,...,m)
1 2 1.0   |
2 1 1.0   |
2 3 1.0   |
1.0E+20   infinity
1.0       default value for entries in c_l
0         # non default entries in c_l
1.0E+20   default value for entries in c_u
0         # non default entries in c_u
0.0       default value for entries in l
0         # non default entries in l
1.0       default value for entries in u
1         # non default entries in u
2 2.0     1 line of non-default indices and values in u
0         default variable type is continuous
1         # non default variable types
3 1       variable 3 is integer
1.0       default value for initial values for x
0         # non default entries in x
0.0       default value for initial values for y
0         # non default entries in y
0.0       default value for initial values for z
0         # non default entries in z
0         # non default names for variables
0         # non default names for constraints
\end{verbatim}
}
